\AliusSetIssueNumber{5}
\AliusSetYear{2021}
\AliusSetPieceSlug{canna-seligman}
\AliusSetPieceType{interview}
\AliusSetTitle{How culture breaks down the mind-body divide}
\AliusSetSubtitle{An interview with Maddalena Canna and Rebecca Seligman}
\AliusSetCreditLine{by Matthieu Koroma}
\AliusSetRunningTitle{Canna \& Seligman ? How culture breaks down the mind-body divide}
\AliusSetCitationLabel{Cite as:}
\AliusSetStandaloneCitation{Canna, M., Seligman, R. \& Koroma, M. (2021). How culture breaks down the mind-body divide. An interview with Maddalena Canna and Rebecca Seligman by Matthieu Koroma. ALIUS Bulletin, 5, 12-27, https://doi.org/10.34700/mfzk-ve81}
\AliusSetAbstract{In this interview, anthropologists Maddalena Canna and Rebecca Seligman explain their recent research agenda investigating how culture shapes the relationship between the body and mind. After having studied trance, they shift their focus to Functional Neurological Disorder (FND), an illness in which sensory or motor symptoms do not match any identified physiological causes. Their approach relies on a mix of anthropological and psychophysiological methods to unravel the mechanisms by which cultural expectations influence the perception and meaning of bodily sensations, a phenomenon that they call "interoceptive affordance". The cultural contextualization of psychophysiological processes puts into question our conceptions of the body and mind divide, and what we consider as normal or pathological. By emphasizing how social factors such as gender, race and medical categorization impact the symptomatology of FND, they highlight how anthropological research can lead to a more encompassing and humanized vision of treatment and care that replaces the category of patient with a participatory actor of the scientific investigation of conscious states and their modification.}
\AliusSetKeywordsLabel{keywords:}
\AliusSetKeywords{Functional Neurological Disorder, trance, psychophysiology, interoception, interoceptive affordance, social science, medical ethics}
\AliusSetTitleSize{20}{24}
\AliusSetSubtitleSize{20}{24}
\AliusSetCreditSize{13}{15}
\AliusClearContributors
\AliusAddContributor{Maddalena Canna}{maddalena.canna@northwestern.edu\\Department of Anthropology and Global Health Program,\\Northwestern University, USA}
\AliusAddContributor{Rebecca Seligman}{r-seligman@northwestern.edu\\Department of Anthropology and Global Health Program,\\Northwestern University, USA}
\AliusAddContributor{Matthieu Koroma}{matthieu.koroma@uliege.be\\Physiology of Cognition lab, GIGA Consciousness,\\University of Liege, Belgium}
