\AliusSetIssueNumber{6}
\AliusSetYear{2023}
\AliusSetPieceSlug{gonzalez}
\AliusSetPieceType{interview}
\AliusSetTitle{An interdisciplinary journey into consciousness\\research}
\AliusSetSubtitle{An interview with Juan González}
\AliusSetCreditLine{By Matthieu Koroma}
\AliusSetRunningTitle{González – An interdisciplinary journey into consciousness research}
\AliusSetCitationLabel{Cite as:}
\AliusSetStandaloneCitation{González, J., Koroma, M. (2023). An interdisciplinary journey into consciousness research. An interview with Juan González by Matthieu Koroma, ALIUS Bulletin, 6, https://doi.org/10.5281/zenodo.7674054}
\AliusSetAbstract{In this interview, Pr. González and Dr. Koroma discuss the purpose and the challenges of an interdisciplinary study of consciousness. They do so by first tracing back the history of ALIUS, a 15 years-old international association dedicated to the scientific study of the diversity of consciousness. By discussing how we can incorporate views from multiple disciplines, contexts, and methodologies in our understanding of consciousness, they offer insights into the role of science and philosophy. They further address the mystery of consciousness, the place of first-hand subjective experience and empirical data in modelling consciousness, and how we can navigate among multiple definitions of consciousness and related notions associated with it like sentience or the spiritual realm.}
\AliusSetKeywordsLabel{keywords :}
\AliusSetKeywords{consciousness, interdiscipline, multidimensional models, psychedelics, neurophenomenology, experiential knowledge, sentience}
\AliusSetTitleSize{20}{24}
\AliusSetSubtitleSize{20}{24}
\AliusSetCreditSize{13}{15}
\AliusClearContributors
\AliusAddContributor{Juan González}{jgonzalez@uaem.mx\\Morelos State University,\\Cuernavaca, Mexico}
\AliusAddContributor{Matthieu Koroma}{matthieu.koroma@uliege.be\\University of Liège,\\Liège, Belgium}
