\AliusSetIssueNumber{7}
\AliusSetYear{2024}
\AliusSetPieceSlug{froese}
\AliusSetPieceType{interview}
\AliusSetTitle{The irruption theory: How mind and\\matter\\meet\\along\\the\\route\\to\\consciousness}
\AliusSetSubtitle{An interview with\\{\AliusLatoLight\fontsize{18}{20}\selectfont Tom Froese}}
\AliusSetCreditLine{By Matthieu Koroma}
\AliusSetRunningTitle{Froese – How mind and matter meet along the route to consciousness}
\AliusSetCitationLabel{Citation:}
\AliusSetStandaloneCitation{Froese, T. \& Koroma, M. (2024). The irruption theory: How mind and matter meet along the route to consciousness. An interview with Tom Froese by Matthieu Koroma. ALIUS Bulletin, XX, XX-XX.}
\AliusSetAbstract{In this interview, Tom Froese and Matthieu Koroma address the irruption theory, an account of how mental actions manifest themselves as unexplained variability in brain and bodily activity. We discuss how this approach can be applied to trace back consciousness in (neuro)physiological signatures such as scale free activity and informational complexity. We highlight the study of altered states of consciousness as an opportunity to disentangle the physiological correlates of consciousness from agency. We conclude by addressing how irruption theory provides insights into the therapeutic effects of psychedelics and the importance of crafting an appropriate context to consciously shape agency and behavior.}
\AliusSetKeywordsLabel{Keywords:}
\AliusSetKeywords{irruption theory, agency, free-will, volition, consciousness, psychedelics}
\AliusSetTitleSize{24}{28}
\AliusSetSubtitleSize{14}{18}
\AliusSetCreditSize{14}{16}
\AliusClearContributors
\AliusAddContributor{Tom Froese}{tom.froese@oist.jp\\OIST, Onna, Japan}
\AliusAddContributor{Matthieu Koroma}{matthieu.koroma@uliege.be\\University of Liege,\\Liege, Belgium}
