\AliusSetIssueNumber{5}
\AliusSetYear{2021}
\AliusSetPieceSlug{chadha}
\AliusSetPieceType{interview}
\AliusSetTitle{Indian philosophy and the value of transformative experiences}
\AliusSetSubtitle{An interview with Monima Chadha}
\AliusSetCreditLine{by Edvard Avil?s and Matthieu Koroma}
\AliusSetRunningTitle{Chadha ? Indian philosophy and the value of transformative experiences}
\AliusSetCitationLabel{Cite as:}
\AliusSetStandaloneCitation{Chadha, M., Avil?s, E. \& Koroma, M. (2021). Indian philosophy and the value of transformative experiences. An interview with Monima Chadha by Edvard Avil?s and Matthieu Koroma. ALIUS Bulletin, 5, 29-38, https://doi.org/10.34700/rn10-az29}
\AliusSetAbstract{In this interview, we engage in a cross-cultural discussion about the diversity of consciousness. Indian philosophy can seem quite cryptic and difficult to follow because it is a primary oral tradition. However, Monima Chadha has developed a series of work aiming at introducing the rich insights of Indian philosophy of mind into Western literature. Indeed, the translation of concepts and practices from the Indian to the Western context requires a solid knowledge of the Indian philosophical tradition in which they form an integrated whole primarily guided by the aim to attain enlightenment or liberation. Despite a careful comparative work of the notions of consciousness developed in both philosophical traditions is still largely lacking, Monima Chadha shares in this interview key insights highlighting how altered states of consciousness can be considered as transformative experiences and meditative practice as an additional resource in the Indian philosophical tradition to gain insight into the nature of consciousness.}
\AliusSetKeywordsLabel{keywords:}
\AliusSetKeywords{Indian philosophy, consciousness, meditation, transformative experiences, psychedelics, no-self}
\AliusSetTitleSize{20}{24}
\AliusSetSubtitleSize{20}{24}
\AliusSetCreditSize{13}{15}
\AliusClearContributors
\AliusAddContributor{Monima Chadha}{monima.chadha@monash.edu\\Department of Philosophy and Monash Centre for Consciousness and Contemplative studies,\\Monash University, Australia}
\AliusAddContributor{Edvard Avil?s}{eda39@cornell.edu\\Sage School of Philosophy,\\Cornell University, USA}
\AliusAddContributor{Matthieu Koroma}{matthieu.koroma@uliege.be\\Physiology of Cognition lab, GIGA Consciousness,\\University of Liege, Belgium}
