\AliusSetIssueNumber{5}
\AliusSetYear{2021}
\AliusSetPieceSlug{olson-yaden}
\AliusSetPieceType{conversation}
\AliusSetTitle{Are the Subjective Effects of Psychedelics\\Necessary for Their Enduring Therapeutic\\Effects?}
\AliusSetSubtitle{A conversation with David E. Olson and David\\B. Yaden}
\AliusSetCreditLine{Hosted by George Fejer}
\AliusSetRunningTitle{David Olson \& David Yaden – Subjective and Therapeutic Effects of Psychedelics}
\AliusSetCitationLabel{Cite as:}
\AliusSetStandaloneCitation{Olson, D.E., Yaden, D.B., Fejer, G. (2021). Are the Subjective Effects of Psychedelics Necessary for Their Enduring Therapeutic Effects? A conversation with David E. Olson and David B. Yaden hosted by George Fejer, ALIUS Bulletin, 5, 40-57, https://doi.org/10.34700/0v21-5n82}
\AliusSetAbstract{ALIUS recently invited Dr. David E. Olson and Dr. David B. Yaden to discuss whether or not the subjective effects of psychedelics are necessary for their enduring therapeutic benefit in an interactive online discussion. The aim of this discussion is to examine their most recent back-to-back publication wherein Yaden \& Griffiths (2021) emphasized the subjective effects of psychedelics, such as mystical experiences, in relation to their long-lasting therapeutic effects, whereas Olson (2021) emphasized that there are certain therapeutic benefits related to the psychoplastogenic properties of these substances that are unrelated to their subjective effects. We invited the authors to clarify the fine points of their arguments and to tease out any point of disagreement between these two perspectives. What follows is an edited transcript of their discussion.}
\AliusSetKeywordsLabel{keywords:}
\AliusSetKeywords{psychedelics, psychoplastogens, subjective effects, therapeutics benefits, treatment ethics}
\AliusSetTitleSize{20}{24}
\AliusSetSubtitleSize{20}{24}
\AliusSetCreditSize{13}{15}
\AliusClearContributors
\AliusAddContributor{David E. Olson}{deolson@ucdavis.edu\\Department of Chemistry,\\Department of Biochemistry \& Molecular Medicine,\\Center for Neuroscience,\\UC Davis, USA}
\AliusAddContributor{David B. Yaden}{dyaden1@jhmi.edu\\Center for Psychedelic \& Consciousness Research,\\Johns Hopkins University, USA}
\AliusAddContributor{George Fejer}{gyorgyf@live.com\\Team coordinator of the\\ALIUS research network}
